---
title: "Terrain derivatives"
subtitle: "STEC project"
author: "Lorena Abad"
date: "February 9, 2021"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---

## Setup
Call libraries, data and initialize SAGA

```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, out.width = "100%")

```
```{r Libraries}
library(here)
# remotes::install_github('loreabad6/terrain')
library(terrain)
library(terra)

```
```{r Directories}
stec_dir = "R:/RESEARCH/02_PROJECTS/01_P_330001/119_STEC/04_Data/Gullies-Mangatu"
```

Set a prefix for file naming

```{r prefix, eval = T}
prefix = "mangatu_"
in_dem = here(stec_dir, 'Mangatu_feature_extraction', 'Mangatu_LiDAR_DEM_2019.tif')
out = here('data_rs', 'terrain')
```

### Convert to SAGA format

```{r Load, eval = F}
elev_to_sgrd(
  elev_filename = in_dem,
  out_filename = here(out, paste0(prefix, "dem.sgrd"))
)
```

### Initialize SAGA
To create the terrain derivatives we need to have a SAGA local GUI

```{r init, eval = F}
saga_path = here('..', 'Earthflows_R/software/saga-7.6.1_x64')
saga_env = init_saga(saga_path)
```

## Calculate derivatives

```{r demset, eval = F}
dem = here(out, paste0(prefix, 'dem.sgrd'))

```
```{r calc, eval = F}
elev_to_morphometry(
  elev_sgrd = dem,
  out_dir = out,
  prefix = prefix,
  envir = saga_env,
  cplan = T,
  cprof = T,
  ddgrd = T,
  mbidx = T,
  slope = T,
  textu = T,
  tridx = T
)

elev_to_lighting(
  elev_sgrd = dem,
  out_dir = out,
  prefix = prefix,
  envir = saga_env,
  shade = T
)

elev_to_terrain_analysis(
  elev_sgrd = dem,
  out_dir = out,
  prefix = prefix,
  envir = saga_env,
  lsfct = T,
  flow = T,
  spcar = T,
  twidx = T
)

slope = here(out, paste0(prefix, 'slope.sgrd'))
spcar = here(out, paste0(prefix, 'spcar.sgrd'))
flow = here(out, paste0(prefix, 'flow.sgrd'))

elev_to_hidrology(
  elev_sgrd = dem,
  slope_sgrd = slope,
  spcar_sgrd = spcar,
  out_dir = out,
  prefix = prefix,
  envir = saga_env,
  spidx = T
)

elev_to_channel(
  elev_sgrd = dem,
  flow_sgrd = flow,
  out_dir = out,
  prefix = prefix,
  envir = saga_env,
  chnet = TRUE,
  vdcnw = TRUE,
  init_value = 500,
  chnet_shp = TRUE
)
```

### Convert from SAGA to GeoTIFF
The current version of GDAL (3.1.4) offered by OSGEO4W
does not seem to support WKT crs inputs. Hence, we need
to rely still on the ArcGIS toolbox to reproject the data. 

```{r translate, eval = F}
c = rgdal::showWKT("init=epsg:2193")
terrain_to_tif(
  out
  # out_crs = c
)
```

## Remarks:
Some layers still need to be checked visually!

```{r render, eval = F, include = F}
o = knitr::spin('pre-processing/terrain_derivatives.R', knit = FALSE)
rmarkdown::render(o)
```

